Synthesizing Quotient Lenses

### 2018-03-13 papers/QuotientLenses/main.tex
in progress edits to section 4

\section{Quotient Regular Expressions}
\label{QRE}
A Quotient Regular Expression (or QRE) is a regular expression $R$ augmented
with syntax that expresses an equivalence relation on the language of $R$. This
section describes in further detail the set of QRE combinators that we described
in \cref{sec:example}.
\iflastminute 
\saz{Why does the ``cref'' macro not capitalize Section?}
\bcp{Dunno.  Because SIGPLAN does not want it to??  But maybe we should be
  using autoref instead anyway...}
\fi

\subsection{Syntax of Semantics of QREs}
Formally, the language of Quotient Regular Expressions (QREs) is given by the
following grammar:
\newcommand{\bsep}{\ \ \sep{} \ \ }
\begin{align*}
Q := \; & \normalize{R}{R'}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
 | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
\sep Q_2) \bsep Q^*,
\end{align*}
where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
functions between regular languages, and $s$ ranges over character strings.

Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
by the regular expression $R$, each QRE $Q$ enables us to express four semantic objects

\begin{tabular}{rl}
  $W(Q)$  & a regular expression (the ``whole'' of $Q$),\\
  $\eqrel{Q}$  & an equivalence relation on $\mathcal{L}(W(Q))$,\\
  $K(Q)$ & a regular expression (the ``kernel'' of $Q$) such that
           $\mathcal{L}(K(Q))$ forms a complete \\
             & set of representatives for $\eqrel{Q}$, and \\
  $\canonize(Q)$ & 
  a ``canonizing'' function that, given any $w \in \mathcal{L}(W(Q))$,
  yields $\canonize(Q)(w)$ \\ & as the unique $k$ in $\mathcal{L}(K(Q))$ such
  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
\end{tabular}
  % of type $\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$  which

The well-formedness constraints for QREs ensure that these four semantic
components fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
equivalence classes are determined by \canonize(Q).

@@

The relationship among the semantic objects of a QRE can be understood in terms
of the \normalize{R}{S}{f} combinator, which exposes each of these pieces
explicitly.  Its whole language is just $R$, its kernel language is just
$S$, and $\canonize(\normalize{R}{S}{f}) = f$.  Its equivalence relation is
determined by the fibres of $f$, so we have
$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$.

@@

\begin{prooftree}
\AxiomC{$\mathcal{L}(S) \subseteq \mathcal{L}(R)$}
\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$}
\AxiomC{$f$ is surjective}
\AxiomC{$f = f^2$}
\RightLabel{(Normalize)}
\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
\end{prooftree}

Semantically, the \kw{normalize} QRE is universal---each of the other
combinators $Q$ is equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some
surjective, idempotent function
$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.   However,
although \kw{normalize} is powerful, verifying that a canonization function $f$
is surjective and idempotent is in general undecidable.  This means that a
programmer wishing to use \kw{normalize} must discharge strong proof
obligations, which is cumbersome in practice.\footnote{In our implementation we allow a programmer to use \kw{normalize} at their
own risk without checking these side conditions as an ``escape hatch'' for the
case when other QRE combinators are insufficient.\bcp{Do we use this in any
  of the programs we've written?}}

The remaining QRE combinators provide simpler, more compositional ways of
building canonizers.  Nevertheless, the \kw{normalize} combinator gives a
sufficient condition for potential QREs to be well-behaved, particularly with
respect to the strategy for synthesizing quotient lenses which we shall
introduce in Section~\ref{synth}, so we use it to guide the design of our QRE
combinators, as we see next.

@@

Figure~\ref{fig:wk} gives the inductive definitions of the whole and
kernel languages $W(Q)$ and $K(Q)$ for all of the QRE forms.
\begin{figure}[t]
\centering
\[
\begin{array}{l@{\quad}l@{\quad}l}

Q & W(Q) & K(Q) \\ \hline
R & R & R \\
\project{R}{s} & R & s \\
\squash{Q}{Q_1}{f} & W(Q) \sep W(Q_1) & K(Q_1) \\
\normalize{R}{R'}{f} & R & R' \\
Q_1 \; ; \; Q_2 & W(Q_1) & K(Q_2) \\
Q_1 \cdot Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
Q_1 \sep Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
Q^* & W(Q)^* & K(Q)^* \\
\end{array}
\]
\[
\begin{array}{r@{\quad}l}
W( \perm{Q_1, \ldots, Q_n}{Q} ) = &
\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
W(Q) \cdot W(c_{\sigma(n)})\\
K( \perm{Q_1, \ldots, Q_n}{Q} ) = & K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q)
\cdot K(Q_n)
\end{array}
\]
\caption{Whole and Kernel Regular Expressions}
\label{fig:wk}
\end{figure}
The \kw{squash} and \kw{permutation} combinators have the two most
interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \kw{squash} combinator
takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
and a function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
canonizes $W(Q_2)$ into $K(Q_2)$ using $\canonize(Q_2)$.

For the $\perm{Q_1, \ldots, Q_n}{Q}$ combinator, the whole language is the
union of languages of the form $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
W(Q) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
$S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
the \kw{permutation} combinator allows for the string to match any
permutation the $Q_i$'s while also taking into account the separator $Q$ in
between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
is the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$,
because the canonical permutation is the identity permuation, with each of the
parts of the string that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
respectively.

Figure~\ref{fig:canonizers} gives the inductive definitions of the
$\canonize{}$ function for each QRE.
\begin{figure}[t]
\begin{center}
\[
\begin{array}{l@\quad l @\quad l}
\canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
\canonize(\project{R}{s})(w) &=& s\\
\canonize(\squash{Q_1}{Q_2}{f})(w) &=&
\begin{cases}
\canonize(Q_1)(f(w)) & \text{if } w \in \mathcal{L}(W(Q_1))\\
\canonize(Q_2)(w) & \text{otherwise}
\end{cases}\\
\canonize(\normalize{R}{R'}{f}) &=& f\\
\canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\
\canonize(Q_1 \sep Q_2)(w) &=&
\begin{cases}
\canonize(Q_1)(w) & \text{if } w \in \mathcal{L}(W(Q_1))\\
\canonize(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
\end{cases}\\
\canonize(Q^*) &=& \canonize(Q)^* \\
\end{array}
\]
\end{center}
\begin{align*}
&\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
&= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
(\canonize(Q_n)(r_n))
\end{align*}
\caption{QRE canonizers}
\label{fig:canonizers}
\end{figure}
Again, the \kw{permutation} combinator gives the most interesting
definition,
\begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
(\canonize(Q_n)(r_n))
\end{align*}
\noindent which, as we have explained, is defined so that the strings that match
$Q_i, Q$ are placed in the canonical permutation which is the identity
permutation before $\canonize(Q_i)$ and $\canonize(Q)$ are applied
respectively.

Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
equivalence relation $\eqrel{Q}$. This is the set-theoretic semantics of QREs
as an equivalence relation on a regular language.

\begin{figure}[t]
\centering
\[
\begin{array}{l@{}l@{\ \ }l}
w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
f(w)=f(w') \\
w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
\text{ or } w \equiv_{Q_2} w' \\
w \; \equiv_{Q_1 \; ; \; Q_2} \; w' &\iff& \exists k, k' \in
\mathcal{L}(K(Q_2)) \text{ such that } w \; \equiv_{Q_1} \; k, \; w' \;
\equiv_{Q_1} \; k', \text{ and } k \; \equiv_{Q_2} \; k'\\
w \; \equiv_{Q_1 \cdot Q_2} \; w'  &\iff& w = r_1
\cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{Q_1}
\; {r'}_1, \; r_2 \; \equiv{Q_n} \; {r'}_2\\
w \; \equiv_{Q_1 \sep Q_2} \; w' &\iff& w \; \equiv_{Q_1} \; w'
\text{ or } \; w \; \equiv_{Q_2} \; w'\\
w \; \equiv_{Q^*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
= {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{Q} \; {r'}_i
\\
w \; \equiv_{\perm{Q_1, \ldots, Q_n}{Q}} \; w' &\iff& w = r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
\cdot {r'}_{\theta(n)} \\
& & \text{for some } \sigma, \theta \in S_n, \text{ with } r_i \;
\equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{Q} \; s'_{k}
\end{array}
\]
\caption{QRE Equivalence Relations}
\label{fig:relations}
\end{figure}
\subsection{Ambiguity and Well Formed QREs}
\label{subsec:well-formed-qres}
An issue that we address before presenting the full set of inference
rules for QREs is the issue of {\em ambiguity}, which appears in the
premises of the inference rules for the regular combinators. If $Q_1$ and $Q_2$
are QREs, then when applying the regular combinators to QREs, we require that
if a string $s$ matches any of the regular expressions $W(Q_1) \cdot W(Q_2)$,
\quad $W(Q_1) \sep W(Q_2)$, \quad $W(Q_1)^*$, \quad $K(Q_1) \cdot K(Q_2)$,
\quad $K(Q_1) \sep K(Q_2)$, \quad $K(Q_1)^*$, then $s$ matches that regular
expression in only one way.

>>>>

\section{Quotient Regular Expressions}
\label{QRE}
A Quotient Regular Expression (or QRE) is a regular expression $R$ augmented
with syntax that expresses an equivalence relation on the language of $R$. This
section formalizes the set of QRE combinators that we introduced informally
in \cref{sec:example}.
\iflastminute 
\saz{Why does the ``cref'' macro not capitalize Section?}
\bcp{Dunno.  Because SIGPLAN does not want it to??  But maybe we should be
  using autoref instead anyway...}
\fi

\subsection{Syntax and Semantics of QREs}
Formally, the language of Quotient Regular Expressions (QREs) is given by the
following grammar:
\newcommand{\bsep}{\ \ \sep{} \ \ }
\begin{align*}
Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
 | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
\sep Q_2) \bsep Q^*
\end{align*}
where $Q$ ranges over QREs, 
$R$ ranges over regular expressions, 
$f$ ranges over functions between regular languages, and 
$s$ ranges over character strings.

Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
by the regular expression $R$, each QRE $Q$ yields four semantic
objects:

\begin{tabular}{rl}
  $W(Q)$       & A regular expression, denoting the ``whole'' of $Q$\\
  $\eqrel{Q}$  & An equivalence relation on $\mathcal{L}(W(Q))$\\
  $K(Q)$       & A regular expression, denoting the ``kernel'' of $Q$, such that
                  $\mathcal{L}(K(Q))$ forms a \\
               & complete set of representatives for $\eqrel{Q}$ \\
  $\canonize(Q)$ & 
  A ``canonizing'' function. Given any $w \in \mathcal{L}(W(Q))$,
  $\canonize(Q)(w)$ \\ & is the unique $k$ in $\mathcal{L}(K(Q))$ such
  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
\end{tabular}
  % of type $\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$  which

\noindent
Intuitivley, $W(Q)$ is the regular expression representing the
external format, while $K(Q)$ is the regular expression representing
the internal format.  The equivalence relation $\eqrel{Q}$ groups
together elements in the language of $W(Q)$ that contain the same
essential information.  The function $\canonize(Q)$ picks the
representative element from each of the resulting equivalence classes.

The well-formedness constraints for QREs ensure that these four semantic
objects fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
equivalence classes are determined by \canonize(Q).

@@

The relationship among the semantic objects of a QRE can be understood in terms
of the combinator \normalize{R_1}{R_2}{f}, which expresses each of these pieces
explicitly.  Its whole language is just $R_1$, its kernel language is just
$R_2$, and its canonizer is just $f$; 
its equivalence relation $\equiv$ is
determined by the fibres of $f$, so we have
$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$ for $s_1$ and $s_2$
in $\mathcal{L}(R_1)$.

@@

\begin{prooftree}
\AxiomC{$\mathcal{L}(R_2) \subseteq \mathcal{L}(R_1)$}
\AxiomC{$f : \mathcal{L}(R_1) \longrightarrow \mathcal{L}(R_2)$}
\AxiomC{$f$ is surjective}
\AxiomC{$f = f^2$}
\RightLabel{(Normalize)}
\QuaternaryInfC{$\wf{\normalize{R_1}{R_2}{f}}$}
\end{prooftree}

Semantically, the \kw{normalize} QRE is universal---each of the other
combinators $Q$ is equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some
surjective, idempotent function 
$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.  
However,
verifying that a canonization
function $f$ is surjective and idempotent is in general undecidable.
Consequently, a programmer wishing to use \kw{normalize} must
discharge strong proof obligations, which is cumbersome in
practice.\footnote{In our implementation we allow a programmer to
use \kw{normalize} at their own risk without checking these side
conditions as an ``escape hatch'' for the case when other QRE
combinators are insufficient.\bcp{Do we use this in any of the
programs we've written?}}

The remaining QRE combinators, which we discuss next, provide simpler,
more compositional ways of building canonizers.  Nevertheless,
the \kw{normalize} combinator provides a useful guide in the design of
these combinators because it gives a sufficient condition for the
well-formedness of any potential QREs.

% particularly with respect to the strategy for synthesizing quotient
% lenses (See Section~\ref{synth}), so we use it to guide the design of
% our QRE remaining combinators, which we discuss next.

@@

Figure~\ref{fig:wk} gives the inductive definitions of the whole and
kernel languages $W(Q)$ and $K(Q)$ for all of the QRE combinators.
\begin{figure}[t]
\centering
\[
\begin{array}{l@{\quad\quad}l@{\quad\quad}l}

Q & W(Q) & K(Q) \\ \hline
R & R & R \\
\project{R}{s} & R & s \\
\squash{Q_1}{Q_2}{f} & W(Q_1) \sep W(Q_2) & K(Q_2) \\
\normalize{R_1}{R_2}{f} & R_1 & R_2 \\
Q_1 \; ; \; Q_2 & W(Q_1) & K(Q_2) \\
Q_1 \cdot\; Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
Q_1 \sep \; Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
Q^* & W(Q)^* & K(Q)^* \\
\end{array}
\]
\[
\begin{array}{r@{\quad}l}
W( \perm{Q_1, \ldots, Q_n}{Q} ) = &
\bigcup \limits_{\sigma \in S_n} W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
W(Q) \cdot W(Q_{\sigma(n)})\\
K( \perm{Q_1, \ldots, Q_n}{Q} ) = & K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q)
\cdot K(Q_n)
\end{array}
\]
\caption{Whole and Kernel Regular Expressions for QRE combinators. 
In describing regular expressions, we use the notations $\sep$ and $\bigcup$ for
alternation, the notation $\cdot$ for concatenation, and the notation
$*$ for Kleene closure.  
We use the notation $S_n$ to denote the set of all permutations of the
numbers $1$ to $n$.
}
\label{fig:wk}
\end{figure}
The \kw{squash} and \kw{permutation} combinators have the two most
interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
of $Q$ is $W(Q_1) \sep W(Q_2)$ because the \kw{squash} combinator
merges the whole language $W(Q_1)$ of $Q_1$ with the whole language $W(Q_2)$ of $Q_2$.
The function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
canonizes $W(Q_2)$ into $K(Q_2)$ using $\canonize(Q_2)$.

For the $\perm{Q_1, \ldots, Q_n}{Q}$ combinator, the whole language is the
union of languages of the form $W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
W(Q) \cdot W(Q_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
$S_n$ is the set of all permutations of the numbers $1$ to $n$. Intuitively,
the \kw{permutation} combinator allows for the string to match any
permutation of the $Q_i$'s while preserving the separator $Q$ in
between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
is the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$,
because the canonical permutation is the identity permuation, with each of the
parts of the input that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
respectively.

Figure~\ref{fig:canonizers} gives the inductive definitions of the
$\canonize{}$ function for each QRE.
\begin{figure}[t]
\begin{center}
\[
\begin{array}{l@\quad l @\quad l}
\canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
\canonize(\project{R}{s})(w) &=& s\\[1ex]
\canonize(\squash{Q_1}{Q_2}{f})(w) &=&
\begin{cases}
\canonize(Q_1)(f(w)) & \text{if } w \in \mathcal{L}(W(Q_1))\\
\canonize(Q_2)(w) & \text{otherwise}
\end{cases}\\[3ex]
\canonize(\normalize{R_1}{R_2}{f}) &=& f\\
\canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\[1ex]
\canonize(Q_1 \sep Q_2)(w) &=&
\begin{cases}
\canonize(Q_1)(w) & \text{if } w \in \mathcal{L}(W(Q_1))\\
\canonize(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
\end{cases}\\[3ex]
\canonize(Q^*) &=& \canonize(Q)^* \\
\end{array}
\]
\end{center}
\begin{align*}
&\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
&= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
(\canonize(Q)(s_{n-1}))\cdot
(\canonize(Q_n)(r_n))
\end{align*}
\caption{QRE canonizers. \ksf{Solomon: please define/explain the . and *
operators on functions.}}
\label{fig:canonizers}
\end{figure}
The \kw{permutation} combinator gives rise to the most interesting
definition:
\begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
(\canonize(Q)(s_{n-1}))\cdot
(\canonize(Q_n)(r_n))
\end{align*}
\noindent 
which places the strings that match $Q_i$ and $Q$ according to the
canonical permutation (i.e., the identity permutation) before applying
the canonizers $\canonize(Q_i)$ and $\canonize(Q)$, respectively.

Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
equivalence relation $\eqrel{Q}$, which is the set-theoretic semantics
of a QRE $Q$ as an equivalence relation on the regular language $W(Q)$.

\begin{figure}[t]
\centering
\[
\begin{array}{l@{}l@{\ \ }l}
w \; \equiv_{\normalize{R_1}{R_2}{f}} \; w' &\iff&
f(w)=f(w') \\
w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
\text{ or } w \equiv_{Q_2} w' \\
w \; \equiv_{Q_1 \; ; \; Q_2} \; w' &\iff& \exists k, k' \in
\mathcal{L}(K(Q_2)) \text{ such that } w \; \equiv_{Q_1} \; k, \; w' \;
\equiv_{Q_1} \; k', \text{ and } k \; \equiv_{Q_2} \; k'\\
w \; \equiv_{Q_1 \cdot Q_2} \; w'  &\iff& w = r_1
\cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{Q_1}
\; {r'}_1, \; r_2 \; \equiv{Q_n} \; {r'}_2\\
w \; \equiv_{Q_1 \sep Q_2} \; w' &\iff& w \; \equiv_{Q_1} \; w'
\text{ or } \; w \; \equiv_{Q_2} \; w'\\
w \; \equiv_{Q^*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
= {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{Q} \; {r'}_i
\\
w \; \equiv_{\perm{Q_1, \ldots, Q_n}{Q}} \; w' &\iff& w = r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
\cdot {r'}_{\theta(n)} \\
& & \text{for some } \sigma, \theta \in S_n, \text{ with } r_i \;
\equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{Q} \; s'_{k}
\end{array}
\]
\caption{QRE Equivalence Relations}
\label{fig:relations}
\end{figure}
\subsection{Ambiguity and Well Formed QREs}
\label{subsec:well-formed-qres}
An issue that we address before presenting the full set of inference
rules for QREs is the issue of {\em ambiguity}, which appears in the
premises of the inference rules for the regular combinators. If $Q_1$ and $Q_2$
are QREs, then when applying the regular combinators to QREs, we require that
if a string $s$ matches any of the regular expressions $W(Q_1) \cdot W(Q_2)$,
\quad $W(Q_1) \sep W(Q_2)$, \quad $W(Q_1)^*$, \quad $K(Q_1) \cdot K(Q_2)$,
\quad $K(Q_1) \sep K(Q_2)$, \quad $K(Q_1)^*$, then $s$ matches that regular
expression in only one way.
